```{r config, echo=FALSE, message=FALSE, warning=FALSE, results="hide"}
# Run config first to use dynamic variables in the YAML
if (!require("upstartr")) install.packages("upstartr")
library("upstartr")
run_config()
```

---
title: "`r getOption("startr.title")`"
author: "`r getOption("startr.author")`"
date: "`r format(Sys.Date(), "%B %d, %Y")`"
output:
  html_document:
    toc: true
    number_sections: true    
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: false    
    highlight: textmate
    theme: cosmo
    df_print: kable
    self_contained: true
    code_folding: hide
---

```{r elt, echo=FALSE, message=FALSE, warning=FALSE, fold_source=TRUE}

# source(dir_src("theme.R"))
# source(dir_src("palette.R"))
run_process()
run_analyze()
run_visualize()
```



## Take-aways


### City of Winnipeg salary disclosure docs, 2016-2020
**Across City of Winnipeg departments, the Winnipeg Police Service ranks in first place for:**

-   has the most staff on Winnipeg's salary disclosure list for 2016-2020 (1,381 - 1,853 people)
-   has the highest sum of total compensation for 2016-2020, based on disclosure list alone (\$168,865,173 - \$192,917,553)
-   recorded the highest compensation in the city in the years 2016, 2019 and 2020 ($368,883, $272,876, $291,834 respectively). 
    -   The Chief Administrative Office reported the highest compensation for 2017 and 2018 ($261,868 and $266,483).
    -   The officer who earned \$368,883 in 2016 appears to have left the force that year. Could potentially semi-trace their salary pre-2016. TBD



------------------------------------------------------------------------

### Data tables

#### winnipeg_all_salary_disclosure_2016_2020_grouped_ranked

```{r dt_winnipeg_all_salary_disclosure_2016_2020_grouped_ranked, echo=TRUE, message=TRUE, warning=TRUE }

datatable(winnipeg_all_salary_disclosure_2016_2020_grouped_ranked,
  caption = "Table: winnipeg_all_salary_disclosure_2016_2020_grouped_ranked",
  class = "cell-border stripe",
  rownames = FALSE,
  filter = "top",
  options = list(
    pageLength = 10,
    order = list(list(3, "asc"), list(1, "asc")),
    language = list(search = 'Filter:')
  )
)
```

#### winnipeg_all_salary_disclosure_2016_2020_grouped

```{r dt_winnipeg_all_salary_disclosure_2016_2020_grouped, echo=TRUE, message=TRUE, warning=TRUE }

datatable(winnipeg_all_salary_disclosure_2016_2020_grouped,
  caption = "Table: winnipeg_all_salary_disclosure_2016_2020_grouped",
  class = "cell-border stripe",
  rownames = FALSE,
  filter = "top",
  options = list(
    pageLength = 10,
    order = list(list(1, "desc")),
    language = list(search = 'Filter:'),
    search = list(regex = TRUE, caseInsensitive = FALSE, search = 'Police')
  )
)
```

------------------------------------------------------------------------

#### winnipeg_police_salary_disclosure_2016_2020

Winnipeg Police Service staff are named but officers (Constables, etc.) are not.

```{r dt_winnipeg_police_salary_disclosure_2016_2020, echo=TRUE, message=TRUE, warning=TRUE }

datatable(winnipeg_police_salary_disclosure_2016_2020,
  caption = "Table: winnipeg_police_salary_disclosure_2016_2020",
  class = "cell-border stripe",
  rownames = FALSE,
  filter = "top",
  options = list(
    pageLength = 25,
    order = list(list(5, "asc"))
  )
)
```


----------

## NOTES and TODO

-   WPS is encoded as both `Police Service` and `Police Services`.

    -   ~~TODO: Normalize department names~~

-   Some entries are salaries below \$70,000

    -   TODO: Set and filter for minimum salary

-   TODO: Confirm if OT is included (across all years?)

